---
title: "Week 6 Code Along Lab"
author: "Katelyn Toigo"
date: "11/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(ggpubr)
library(broom)
```

## A rank-based test example (Mann Whitney U)

Mann-Whitney U to compare ranks (medians) between two unpaired samples (non-parametric alternative to two-sample t-test)

We use set.seed() here to create a “pseudorandom” sample, so that we all get the same samples – otherwise we’d all get something different! We use sample.int() to create random samples with integers from 1 to x, of size = ?

Mann-Whitney U to compare ranks (medians) between two unpaired samples (non-parametric alternative to two-sample t-test)

view groups
view data histograms

```{r}
set.seed(1414)
gp_1 <- sample.int(20, size = 15, replace = TRUE)

set.seed(1424)
gp_2 <- sample.int(30, size = 15, replace = TRUE)

gp_1
gp_2

hist(gp_1)
hist(gp_2)
```

Here, we’ll perform Mann-Whitney U to answer “Is there a significant difference in ranks (medians) between gp_1 and gp_2?” using the wilcox.test() function.

```{r}
my_mwu <- wilcox.test(gp_1, gp_2)

my_mwu
```

What does that p-value of 0.28 actually mean? It means that if the null hypothesis is true (these samples were drawn from populations with the same median), there is a probability of 0.28 that we could have found median values at least as different as ours by chance. In other words: not sufficient evidence to reject the null hypothesis of equal ranks (or medians) using a significance level of 0.05.


## Simple linear regression

Let’s make an exploratory scatterplot of penguing flipper length versus body mass

```{r}
ggplot(data = penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point()
```

We should ask questions about our exploratory visualization, like:

Does it look like a linear relationship makes sense?
Do we have any concerns about modeling as a linear relationship?
Any notable outliers?
Initial thoughts about homoscedasticity (explored more later)?
Here, it looks like overall a linear relationship between flipper length and body mass makes sense here

Once we’ve decided that a linear relationship makes sense, we’ll model it using lm().

```{r}
# Linear model, stored as penguin_lm:
penguin_lm <- lm(body_mass_g ~ flipper_length_mm, data = penguins)

# Return the complete overview:
summary(penguin_lm)
```

Check the Week 6 Part 2 lecture video for information about how to interpret different pieces of this output, but a couple to highlight:

- Both the intercept and flipper_length_mm coefficients are significantly different from zero (not super interesting)

- The Multiple R2 value is 0.759 - meaning that 75.9% of variance in body mass is explained by flipper length


## Access model outputs

We can access the coefficients for the model using:

- The slope is 49.69 (g / mm)

- The y-intercept is -5780.83 (g)

- The full equation is mass = 49.69*(flipper length) + (-5780.83)

But trying to get all of the statistical information from the summary() function would be kind of a mess.

We can use the broom::tidy() function to get the model outputs in nice data frame format:

```{r}
penguin_lm_tidy <- broom::tidy(penguin_lm)

penguin_lm_tidy
```

```{r}
# Get the intercept: 
penguin_int <- penguin_lm_tidy$estimate[1]
penguin_int
```

```{r}
# Then to get the flipper_length coefficient:
penguin_coef <- penguin_lm_tidy$estimate[2]
penguin_coef
```

```{r}
# Metrics at a glance: 
penguin_lm_out <- broom::glance(penguin_lm)
penguin_lm_out
```

We can use the results of both to write a statement about the model that will **automatically update** if anything about the model changes! For example: 

"Simple linear regression was used to explore the relationship between penguin flipper length (mm) and body mass (g) across all three penguin species, and including both male and female penguins. A significant regression model was found ($\beta$ = `r round(penguin_coef,3)`, F(`r penguin_lm_out$df`,`r penguin_lm_out$df.residual`) = `r round(penguin_lm_out$statistic,1)`, p < 0.001) with an R^2^ of `r round(penguin_lm_out$r.squared,3)`."

**Note:** This might seem *really* tedious to write out, but the advantages are worth it. All values will be automatically updated when the model is updated! Reproducible and way less opportunity for human error. Plus, once you have this template statement made, you can reuse it for future regression models and just replace `penguin_lm_out` and `penguin_coef` with the appropriate objects for your new model! 


## Explore model assumptions

Recall that we have assumptions for linear regression we need to explore, some related to the residuals.

- Linearly related variables (CHECK - already looked & thought hard)

- Normally distributed residuals

- Homoscedasticity (constant residuals variance)

- iid residuals (no serial correlation) - more often a concern in time series data

Use the plot() function on the model, which will automatically create four useful visualizations to consider assumptions!

```{r}
plot(penguin_lm)
```

Graphs 1 & 3 are useful for thinking about homoscedasticity; graph 2 (QQ plot) helps us consider normality of residuals; graph 4 reveals the Cook’s distance (a measure of how much leverage any single observation has on the model).


## Visualize the model

Now that we’ve explore the assumptions and have decided that linear regression is a valid tool to describe the relationship between flipper length and body mass, let’s look at the model.

Use geom_smooth(method = "lm") to add a linear model to an existing scatterplot

Use stat_cor() and/or stat_regline_equation() to add equation information directly to the plot panel, at an x- and y-position that you specify

```{r}
ggplot(data = penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point(size = 2) +
  geom_smooth(method = "lm",
              color = "purple",
              size = 0.5,
              fill = "gray10",
              alpha = 0.5) +
  theme_light() +
  ggpubr::stat_regline_equation(label.x = 180, label.y = 5700)
```


## Find Pearson’s r for correlation

The coefficient of determination, R2, which tells us how much of the variance in the dependent variable is explained by the model.

We might also want to explore the strength of the correlation (degree of relationship) between two variables which, for two linearly related continuous variables, can be expressed using Pearson’s r.

Pearson’s r ranges in value from -1 (perfectly negatively correlated - as one variable increases the other decreases) to 1 (perfectly positively correlated - as one variable increases the other increases). A correlation of 0 means that there is no degree of relationship between the two variables.

Typical guidelines look something like this (there’s wiggle room in there):

r = 0: no correlation

r < |0.3|: weak correlation

r between |0.3| and |0.7|: moderate correlation

r > |0.7|: strong correlation

We’ll use the cor.test() function, adding the two vectors (flipper_length_mm and body_mass_g) as the arguments. The function reports the Pearson’s r value, and performs a hypothesis test with null hypothesis that the correlation = 0.

```{r}
penguins_cor <- cor.test(penguins$flipper_length_mm, penguins$body_mass_g)

penguins_cor
```

Here, we see that there is a strong positive correlation between penguin flipper length and body mass (*r* = `r round(penguins_cor$estimate,2)`, t(`r penguins_cor$parameter`) = `r round(penguins_cor$statistic,2)`, p < 0.001). 